---
title: "Testing effect of Floral constancy, cost of switching and unequal reward abundnace on foraging energetics"
output: html_notebook
---


Testing effect of Floral constancy, cost of switching and unequal reward abundnace on foraging energetics

#unequal abundance, no cost of switching

Analysis 1:

First, randomly taking 10000 pollinator visits with CI ranging from 0 to 1
RW: rare moreph with high nectar reward.
GR and GW are common morephs with low rewards.
Here morphs can be replaced by 'Species'.

```{r}
# Define the number of visits
n_visits <- 60

# Define the total number of plants
n_plants <- 1000

# Define the plant species proportions
plant_species_proportions <- c(RW = 0.1, GR = 0.3, GW = 0.6)

# Define the energy gained from each plant species
plant_species_energy <- c(RW = 35, GR = 20, GW = 15)

# Define a function to simulate a foraging bout
simulate_foraging_bout <- function(n_visits, n_plants, plant_species_proportions, CI_range) {
  # Initialize the visit sequence
  visit_sequence <- character(n_visits)
  
  # Initialize the CI value
  CI <- 0
  
  # Simulate the foraging bout
  while (CI < CI_range[1] || CI > CI_range[2]) {
    for (i in seq_len(n_visits)) {
      # Draw a random plant species according to the proportions
      visit_sequence[i] <- sample(names(plant_species_proportions), size = 1, prob = plant_species_proportions * n_plants)
    }
    
    # Count the number of constant and nonconstant transitions
    n_constant_transitions <- sum(visit_sequence[-1] == visit_sequence[-n_visits])
    n_nonconstant_transitions <- sum(visit_sequence[-1] != visit_sequence[-n_visits])
    
    # Calculate the CI value
    CI <- n_constant_transitions / (n_constant_transitions + n_nonconstant_transitions)
  }
  
  # Return the visit sequence and the CI value
  list(visit_sequence = visit_sequence, CI = CI)
}

# Define the number of pollinators
n_pollinators <- 10000

# Initialize the output data frame
output_df <- data.frame(
  pollinator_ID = seq_len(n_pollinators),
  CI = numeric(n_pollinators),
  visits_to_GW = integer(n_pollinators),
  visits_to_GR = integer(n_pollinators),
  visits_to_RW = integer(n_pollinators),
  total_calories_gained = numeric(n_pollinators),
  visit_sequence = I(replicate(n_pollinators, character(n_visits), simplify = FALSE))
)

# Simulate foraging bouts for each pollinator with a CI value in the range of 0.1 to 0.2
for (i in seq_len(n_pollinators)) {
  result <- simulate_foraging_bout(n_visits, n_plants, plant_species_proportions, CI_range = c(0.00001, 1.0))
  
  # Store the result in the output data frame
  output_df$CI[i] <- result$CI
  output_df$visits_to_GW[i] <- sum(result$visit_sequence == "GW")
  output_df$visits_to_GR[i] <- sum(result$visit_sequence == "GR")
  output_df$visits_to_RW[i] <- sum(result$visit_sequence == "RW")
  
  # Calculate and store the total calories gained by this pollinator
  output_df$total_calories_gained[i] <- sum(plant_species_energy[result$visit_sequence])
  
  output_df$visit_sequence[[i]] <- result$visit_sequence
}

# Write the output data frame to a file
#write.csv(output_df, file = "output_nectar_all.csv", row.names = FALSE)


#save output_df as unique dataframe to use it in future
output_nectar_all <- output_df 

```


Plotting
1. Probability of visiting RW vs CI: Graph II
Hypothesis: Probability of finding a rare morph/species should decrease as floral constancy increases.

```{r}
library(ggplot2)
library(extrafont)
library(ggExtra)

p = ggplot(output_nectar_all, aes(x= CI, y=visits_to_RW))+ 
  geom_point()+
  geom_smooth(method=lm, color="black", linetype="dashed", linewidth = 0.5)+
  
  #titles
  xlab("Floral constancy (CI)") + 
  ylab("Number of visits to rare species (A)")+
  
  #axis
  xlim(0,1)+
  

  #Theme
  theme_bw()+
  theme(text=element_text(family="Times New Roman"))
  

p

ggsave("Graph II_Number of visits to rare species vs CI_unequal abundance.png", p, units = "in", width = 6, height = 4.5, dpi = 300)
```

Pearson correlation test
```{r}
res <- cor.test(output_nectar_all$CI, output_nectar_all$visits_to_RW, 
                    method = "pearson")
res
```

2. Total energy gained vs CI
Not needed


```{r}
library(ggplot2)
library(extrafont)

p = ggplot(output_nectar_all, aes(x=CI, y=total_calories_gained))+ 
  geom_point(size = 0.2)+
  geom_smooth(method=lm, color="black", linetype="dashed", linewidth = 0.5)+
  
  #titles
  xlab("Floral constancy (CI)") + 
  ylab("Total reward gained")+

  #axis
  xlim(0,1)+
  
  #Theme
  theme_bw()+
  theme(text=element_text(family="Times New Roman"))

p

ggsave("Graph IV_Floral reward gained vs floral constancy_unequal abundance.png", p, units = "in", width = 8, height = 5, dpi = 300)
```


Pearson correlation test
```{r}
res <- cor.test(output_nectar_all$CI, output_nectar_all$total_calories_gained, 
                    method = "pearson")
res
```


## Adding the cost of switching

Adding cost of switching in this:
switching to RW = 10 units
switching to GR and GW = 5 units

1. With unequal abundance

Codes with five different switching costs:
1/10th, 1/5th, 1/4th, 1/3rd, 1/2nd of their initial floral reward energy


```{r}
# Define the number of visits
n_visits <- 60

# Define the total number of plants
n_plants <- 1000

# Define the plant species proportions
plant_species_proportions <- c(RW = 0.1, GR = 0.3, GW = 0.6)

# Define the energy gained from each plant species
plant_species_energy <- c(RW = 35, GR = 20, GW = 15)

# Define the fractions for the cost of switching
switching_cost_fractions <- c(0,1/10, 1/5, 1/4, 1/3, 1/2)

# Define the number of pollinators
n_pollinators <- 10000

# Loop over each switching cost fraction
for (fraction in switching_cost_fractions) {
  # Define the cost of switching as a fraction of the initial energy values
  switching_cost <- plant_species_energy * fraction

  # Define a function to simulate a foraging bout
  simulate_foraging_bout <- function(n_visits, n_plants, plant_species_proportions, CI_range) {
    # Initialize the visit sequence
    visit_sequence <- character(n_visits)
    
    # Initialize the CI value
    CI <- 0
    
    # Initialize the total energy gained
    total_energy_gained <- 0
    
    # Simulate the foraging bout
    while (CI < CI_range[1] || CI > CI_range[2]) {
      for (i in seq_len(n_visits)) {
        # Draw a random plant species according to the proportions
        visit_sequence[i] <- sample(names(plant_species_proportions), size = 1, prob = plant_species_proportions * n_plants)
        
        # Add the energy gained from this visit, subtracting the switching cost
        total_energy_gained <- total_energy_gained + plant_species_energy[visit_sequence[i]] - ifelse(i > 1 && visit_sequence[i] != visit_sequence[i - 1], switching_cost[visit_sequence[i]], 0)
      }
      
      # Count the number of constant and nonconstant transitions
      n_constant_transitions <- sum(visit_sequence[-1] == visit_sequence[-n_visits])
      n_nonconstant_transitions <- sum(visit_sequence[-1] != visit_sequence[-n_visits])
      
      # Calculate the CI value
      CI <- n_constant_transitions / (n_constant_transitions + n_nonconstant_transitions)
    }
    
    # Return the visit sequence, the CI value, and the total energy gained
    list(visit_sequence = visit_sequence, CI = CI, total_energy_gained = total_energy_gained)
  }

  # Initialize the output data frame
  output_df <- data.frame(
    pollinator_ID = seq_len(n_pollinators),
    CI = numeric(n_pollinators),
    visits_to_GW = integer(n_pollinators),
    visits_to_GR = integer(n_pollinators),
    visits_to_RW = integer(n_pollinators),
    total_calories_gained = numeric(n_pollinators),
    visit_sequence = I(replicate(n_pollinators, character(n_visits), simplify = FALSE))
  )

  # Simulate foraging bouts for each pollinator with a CI value in the range of 0.1 to 0.2
  for (i in seq_len(n_pollinators)) {
    result <- simulate_foraging_bout(n_visits, n_plants, plant_species_proportions, CI_range = c(0.00001, 1.0))
    
    # Store the result in the output data frame
    output_df$CI[i] <- result$CI
    output_df$visits_to_GW[i] <- sum(result$visit_sequence == "GW")
    output_df$visits_to_GR[i] <- sum(result$visit_sequence == "GR")
    output_df$visits_to_RW[i] <- sum(result$visit_sequence == "RW")
    
    # Store the total calories gained by this pollinator
    output_df$total_calories_gained[i] <- result$total_energy_gained
    
    output_df$visit_sequence[[i]] <- result$visit_sequence
  }

  # Save the output data frame in a unique variable
  assign(paste0("output_nectar_all_", fraction), output_df, envir = .GlobalEnv)
}


```



PLotting

doing the ggplot way
```{r}
library(ggplot2)
library(extrafont)
library(gridExtra)
library(dplyr)
library(tidyr)

# Create an empty data frame to store the combined data
combined_df <- data.frame()

# Loop over each switching cost fraction
for (fraction in switching_cost_fractions) {
  # Get the output data frame for this fraction
  output_df <- get(paste0("output_nectar_all_", fraction))
  
  # Add a column for the fraction
  output_df$fraction <- fraction
  
  # Combine the data frames
  combined_df <- rbind(combined_df, output_df)
}

# Reshape the combined data frame to a long format
combined_df_long <- combined_df %>%
  gather(key = "variable", value = "value", -pollinator_ID, -CI, -visit_sequence, -fraction)

#keeping only total_calories_gained
combined_df_long = combined_df_long[combined_df_long$variable == 'total_calories_gained',]

# Create a faceted plot
p <- ggplot(combined_df_long, aes(x=CI, y=value)) + 
  geom_point(size = 0.2) +
  geom_smooth(method=lm, color="black", linetype="dashed", linewidth = 0.5) +
  facet_grid(. ~ fraction)+
  xlab("Floral constancy (CI)") + 
  ylab("Total reward gained") +
  ggtitle("reward gained vs CI for different costs of switching for unequal abd")+
  theme_bw() +
  theme(text=element_text(family="Times New Roman"))

p

ggsave("Graph VI_Floral reward gained vs floral constancy_unequal abundance.png", p, units = "in", width = 10, height = 5, dpi = 300)
```

Running tests
```{r}
# Define the fractions for the cost of switching
switching_cost_fractions <- c(0,1/10, 1/5, 1/4, 1/3, 1/2)

# Loop over each switching cost fraction
for (fraction in switching_cost_fractions) {
  # Get the output data frame for this fraction
  output_df <- get(paste0("output_nectar_all_", fraction))
  
  # Perform the Pearson's correlation test
  res <- cor.test(output_df$CI, output_df$total_calories_gained, method = "pearson")
  
  # Print the result
  print(paste0("Fraction: ", fraction))
  print(res)
}

```


Trying:
```{r}
# Subset the data where CI is approximately 0.5
df_subset <- combined_df[abs(combined_df$CI - 0.5) < 0.01, c(2,5,6,8)]

# Calculate descriptive statistics
summary(df_subset)

# Calculate correlations
cor(df_subset)

# Perform regression analysis
model <- lm(total_calories_gained ~ CI + fraction + visits_to_RW, data = df_subset)
summary(model)

```



#equal abundance but no cost of switching

Same code as the first one, but with equal morph abundance and unequal reward.
```{r}
# Define the number of visits
n_visits <- 60

# Define the total number of plants
n_plants <- 1000

# Define the plant species proportions
plant_species_proportions <- c(RW = 0.33333333, GR = 0.33333333, GW = 0.33333333)

# Define the energy gained from each plant species
plant_species_energy <- c(RW = 35, GR = 20, GW = 15)

# Define a function to simulate a foraging bout
simulate_foraging_bout <- function(n_visits, n_plants, plant_species_proportions, CI_range) {
  # Initialize the visit sequence
  visit_sequence <- character(n_visits)
  
  # Initialize the CI value
  CI <- 0
  
  # Simulate the foraging bout
  while (CI < CI_range[1] || CI > CI_range[2]) {
    for (i in seq_len(n_visits)) {
      # Draw a random plant species according to the proportions
      visit_sequence[i] <- sample(names(plant_species_proportions), size = 1, prob = plant_species_proportions * n_plants)
    }
    
    # Count the number of constant and nonconstant transitions
    n_constant_transitions <- sum(visit_sequence[-1] == visit_sequence[-n_visits])
    n_nonconstant_transitions <- sum(visit_sequence[-1] != visit_sequence[-n_visits])
    
    # Calculate the CI value
    CI <- n_constant_transitions / (n_constant_transitions + n_nonconstant_transitions)
  }
  
  # Return the visit sequence and the CI value
  list(visit_sequence = visit_sequence, CI = CI)
}

# Define the number of pollinators
n_pollinators <- 10000

# Initialize the output data frame
output_df <- data.frame(
  pollinator_ID = seq_len(n_pollinators),
  CI = numeric(n_pollinators),
  visits_to_GW = integer(n_pollinators),
  visits_to_GR = integer(n_pollinators),
  visits_to_RW = integer(n_pollinators),
  total_calories_gained = numeric(n_pollinators),
  visit_sequence = I(replicate(n_pollinators, character(n_visits), simplify = FALSE))
)

# Simulate foraging bouts for each pollinator with a CI value in the range of 0.1 to 0.2
for (i in seq_len(n_pollinators)) {
  result <- simulate_foraging_bout(n_visits, n_plants, plant_species_proportions, CI_range = c(0.00001, 1.0))
  
  # Store the result in the output data frame
  output_df$CI[i] <- result$CI
  output_df$visits_to_GW[i] <- sum(result$visit_sequence == "GW")
  output_df$visits_to_GR[i] <- sum(result$visit_sequence == "GR")
  output_df$visits_to_RW[i] <- sum(result$visit_sequence == "RW")
  
  # Calculate and store the total calories gained by this pollinator
  output_df$total_calories_gained[i] <- sum(plant_species_energy[result$visit_sequence])
  
  output_df$visit_sequence[[i]] <- result$visit_sequence
}

# Write the output data frame to a file
#write.csv(output_df, file = "output_nectar_all.csv", row.names = FALSE)


#save output_df as unique dataframe to use it in future
output_nectar_all <- output_df 

```

Plotting
1. Probability of visiting RW vs CI: Graph II
```{r}
library(ggplot2)
library(extrafont)

p = ggplot(output_nectar_all, aes(x= CI, y=visits_to_RW))+ 
  geom_point()+
  geom_smooth(method=lm, color="black", linetype="dashed", linewidth = 0.5)+
  
  #titles
  xlab("Floral constancy (CI)") + 
  ylab("Number of visits to rare species (A)")+
  
  #axis
  xlim(0,1)+

  #Theme
  theme_bw()+
  theme(text=element_text(family="Times New Roman"))

p

ggsave("Graph I_Number of visits to rare species vs CI_equal abundance.png", p, units = "in", width = 6, height = 4.5, dpi = 300)
```


Pearson correlation test
```{r}
res <- cor.test(output_nectar_all$CI, output_nectar_all$visits_to_RW, 
                    method = "pearson")
res
```

2. Total energy gained vs CI
```{r}
library(ggplot2)
library(extrafont)

p = ggplot(output_nectar_all, aes(x=CI, y=total_calories_gained))+ 
  geom_point(size = 0.2)+
  geom_smooth(method=lm, color="black", linetype="dashed", linewidth = 0.5)+
  
  #titles
  xlab("Floral constancy (CI)") + 
  ylab("Total reward gained")+

  #axis
  xlim(0,1)+
  
  #Theme
  theme_bw()+
  theme(text=element_text(family="Times New Roman"))

p

ggsave("Graph III_Floral reward gained vs floral constancy_equal abundance.png", p, units = "in", width = 8, height = 5, dpi = 300)
```

Pearson correlation test
```{r}
res <- cor.test(output_nectar_all$CI, output_nectar_all$total_calories_gained, 
                    method = "pearson")
res
```


## Adding the cost of switching (equal floral abundnace, unequal reward)

Adding cost of switching in this:
switching to RW = 10 units
switching to GR and GW = 5 units

1. With unequal abundance

Codes with five different switching costs:
0, 1/10th, 1/5th, 1/4th, 1/3rd, 1/2nd of their initial energy


```{r}
# Define the number of visits
n_visits <- 60

# Define the total number of plants
n_plants <- 1000

# Define the plant species proportions
plant_species_proportions <- c(RW = 0.33333333, GR = 0.33333333, GW = 0.33333333)

# Define the energy gained from each plant species
plant_species_energy <- c(RW = 35, GR = 20, GW = 15)

# Define the fractions for the cost of switching
switching_cost_fractions <- c(0, 1/10, 1/5, 1/4, 1/3, 1/2)

# Define the number of pollinators
n_pollinators <- 10000

# Loop over each switching cost fraction
for (fraction in switching_cost_fractions) {
  # Define the cost of switching as a fraction of the initial energy values
  switching_cost <- plant_species_energy * fraction

  # Define a function to simulate a foraging bout
  simulate_foraging_bout <- function(n_visits, n_plants, plant_species_proportions, CI_range) {
    # Initialize the visit sequence
    visit_sequence <- character(n_visits)
    
    # Initialize the CI value
    CI <- 0
    
    # Initialize the total energy gained
    total_energy_gained <- 0
    
    # Simulate the foraging bout
    while (CI < CI_range[1] || CI > CI_range[2]) {
      for (i in seq_len(n_visits)) {
        # Draw a random plant species according to the proportions
        visit_sequence[i] <- sample(names(plant_species_proportions), size = 1, prob = plant_species_proportions * n_plants)
        
        # Add the energy gained from this visit, subtracting the switching cost
        total_energy_gained <- total_energy_gained + plant_species_energy[visit_sequence[i]] - ifelse(i > 1 && visit_sequence[i] != visit_sequence[i - 1], switching_cost[visit_sequence[i]], 0)
      }
      
      # Count the number of constant and nonconstant transitions
      n_constant_transitions <- sum(visit_sequence[-1] == visit_sequence[-n_visits])
      n_nonconstant_transitions <- sum(visit_sequence[-1] != visit_sequence[-n_visits])
      
      # Calculate the CI value
      CI <- n_constant_transitions / (n_constant_transitions + n_nonconstant_transitions)
    }
    
    # Return the visit sequence, the CI value, and the total energy gained
    list(visit_sequence = visit_sequence, CI = CI, total_energy_gained = total_energy_gained)
  }

  # Initialize the output data frame
  output_df <- data.frame(
    pollinator_ID = seq_len(n_pollinators),
    CI = numeric(n_pollinators),
    visits_to_GW = integer(n_pollinators),
    visits_to_GR = integer(n_pollinators),
    visits_to_RW = integer(n_pollinators),
    total_calories_gained = numeric(n_pollinators),
    visit_sequence = I(replicate(n_pollinators, character(n_visits), simplify = FALSE))
  )

  # Simulate foraging bouts for each pollinator with a CI value in the range of 0.1 to 0.2
  for (i in seq_len(n_pollinators)) {
    result <- simulate_foraging_bout(n_visits, n_plants, plant_species_proportions, CI_range = c(0.00001, 1.0))
    
    # Store the result in the output data frame
    output_df$CI[i] <- result$CI
    output_df$visits_to_GW[i] <- sum(result$visit_sequence == "GW")
    output_df$visits_to_GR[i] <- sum(result$visit_sequence == "GR")
    output_df$visits_to_RW[i] <- sum(result$visit_sequence == "RW")
    
    # Store the total calories gained by this pollinator
    output_df$total_calories_gained[i] <- result$total_energy_gained
    
    output_df$visit_sequence[[i]] <- result$visit_sequence
  }

  # Save the output data frame in a unique variable
  assign(paste0("output_nectar_all_", fraction), output_df, envir = .GlobalEnv)
}


```




Plotting with ggplot + for loop
```{r}
library(ggplot2)
library(extrafont)
library(gridExtra)
library(dplyr)
library(tidyr)

# Create an empty data frame to store the combined data
combined_df <- data.frame()

# Loop over each switching cost fraction
for (fraction in switching_cost_fractions) {
  # Get the output data frame for this fraction
  output_df <- get(paste0("output_nectar_all_", fraction))
  
  # Add a column for the fraction
  output_df$fraction <- fraction
  
  # Combine the data frames
  combined_df <- rbind(combined_df, output_df)
}

# Reshape the combined data frame to a long format
combined_df_long <- combined_df %>%
  gather(key = "variable", value = "value", -pollinator_ID, -CI, -visit_sequence, -fraction)

#keeping only total_calories_gained
combined_df_long = combined_df_long[combined_df_long$variable == 'total_calories_gained',]

# Create a faceted plot
p <- ggplot(combined_df_long, aes(x=CI, y=value)) + 
  geom_point(size = 0.2) +
  geom_smooth(method=lm, color="black", linetype="dashed", linewidth = 0.5) +
  facet_grid(. ~ fraction)+
  xlab("Floral constancy (CI)") + 
  ylab("Total reward gained") +
  ggtitle("reward gained vs CI for different costs of switching for equal abd")+
  theme_bw() +
  theme(text=element_text(family="Times New Roman"))

p

ggsave("Graph V_Floral reward gained vs floral constancy_equal abundance.png", p, units = "in", width = 10, height = 5, dpi = 300)
```


Running tests
```{r}
# Define the fractions for the cost of switching
switching_cost_fractions <- c(0, 1/10, 1/5, 1/4, 1/3, 1/2)

# Loop over each switching cost fraction
for (fraction in switching_cost_fractions) {
  # Get the output data frame for this fraction
  output_df <- get(paste0("output_nectar_all_", fraction))
  
  # Perform the Pearson's correlation test
  res <- cor.test(output_df$CI, output_df$total_calories_gained, method = "pearson")
  
  # Print the result
  print(paste0("Fraction: ", fraction))
  print(res)
}

```



